Algebraic Topology Final Report: Morse Theory

### 2013-01-16 paper.tex
submmited version

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath}               % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{amssymb}

@@

\begin{abstract}
  This report covers roughly the part I of J. Milnor's book(citation). The main
  result is the theorem of Morse: every smooth manifold has the same homotopy
  type of a CW-complex. We prove this theorem by developing the concept of
  non-degenerate smooth functions, their index at critical points and the
  correspondence between the index and the dimension of attaching cells. After
  the main theorem, the existence of non-degenerate smooth functions on
  manifolds is established, and as an application, an alternating proof of the
  Lefschetz hyperplane theorem is given.
\end{abstract}

@@

  We may assume $h_{ij}$ is symmetric, since we can write $\bar{h}_{ij} =
  \frac{1}{2}(h_{ij}+h_{ji})$ and get $f = \sum_{i,j}
  x^ix^j\bar{h}_{ij}$. Moreover, $\bar{h}_{i,j}(0) = \frac{1}{2}\frac{\partial^2
    f}{\partial x^i \partial x^j}(0)$, so it is non-singular.

@@

\begin{thm}
  Suppose $f^{-1}[a,b]$ is compact and contains no critical points of $f$. Then
  $M^a$ is diffeomorphic to $M^b$. Furthermore, $M^a$ is a deformation retract
  of $M^b$, so that the inclusion map $\iota : M^a \to M^b$ is a homotopy
  equivalence.
\end{thm}

@@

  For fixed $q \in M$ consider the function $t \mapsto f(\phi_t(q))$. Then If
  $\phi_t(q) \in f^{-1}[a,b]$, then
  \begin{equation}
    \frac{d f(\phi_t(q))}{dt} = X_{\phi_t q}(f) = +1
  \end{equation}
  Hence the correspondence $t \mapsto f(\phi_t(q))$ is linear as long as
  $f(\phi_t(q)) \in [a,b]$. It follows that $f(\phi_{t}(q)) = t + f(q)$ in a
  small neighborhood of $0 \in \RR$, where $q \in f^{-1}[a,b]$.

  Consider the diffeomorphism $\phi_{b-a} : M \to M$. By above it's clear
  that $\phi_{b-a}$ maps $M^a$ diffeomorphically onto $M^b$. This prove the
  first part of the theorem.

@@

\begin{thm}
  Let $f: M \to \RR$ be a smooth function and let $p$ be a non-degenerate
  critical point with index $\lambda$. Setting $f(p) = c$, suppose that $p$ is
  the only critical point in $f^{-1}[c-\epsilon, c+\epsilon]$, for some
  $\epsilon > 0$. Then for all $\epsilon$ sufficiently small, the set
  $M^{c-\epsilon}$ has the same homotopy type of $M^{c+\epsilon}$ with a
  $\lambda$-cell attached.
\end{thm}

@@

For example, we can define $\mu$ as follows. Let $\mu = \epsilon'
\frac{\int_x^{2\epsilon} \alpha(t)\alpha(2\epsilon-t)}dt{\int_0^{2\epsilon}
  \alpha(t)\alpha(2\epsilon-t)}dt$, where $2\epsilon > \epsilon' > \epsilon$ and
\begin{equation}
  \alpha(x) =
  \begin{cases}
    0 & x \le 0 \\
    e^{-1/x} & x > 0
  \end{cases}
\end{equation}
is smooth. Then $0 \ge \mu'(r) > -\frac{\epsilon'}{2\epsilon} > -1$.

@@

\begin{proof}
  \begin{equation}
    \nonumber
    \begin{aligned}
      F &= c - \xi + \eta - \mu(\xi+2\eta) \\
      \frac{\partial F}{\partial \xi} &= -1 - \mu'(\xi+2\eta) < 0 \\
      \frac{\partial F}{\partial \eta} &= 1 - 2\mu'(\xi+2\eta) ge 1 \\
      dF &= \frac{\partial F}{\partial \xi}d\xi+\frac{\partial F}{\partial
        \eta}d\eta
    \end{aligned}
  \end{equation}
  Since the 1-forms $d\xi$ and $d\eta$ are simultaneously 0 only when both
  $\xi$ and $\eta$ equal 0, it follows that $F$ has no critical point in $U$
  other than $p$.
\end{proof}

Now, since $F \le f$, if $q \in F^{-1}[c-\epsilon, c+\epsilon]$, then
$c-\epsilon \le f(q)$. But $q \in F^{-1}(-\infty, c+\epsilon] = f^{-1}(-\infty,
c+\epsilon]$, $f(q) \le c+ \epsilon$. Therefore $q \in f^{-1}(-\infty,
c+\epsilon]$. This proves that $F^{-1}[c-\epsilon, c+\epsilon] \subset
f^{-1}[c-\epsilon, c+\epsilon]$. Being a closed subset of a compact set in a
locally compact Hausdorff space, $F^{-1}[c-\epsilon, c+\epsilon]$ is compact.
Note that $F(p) = c- \mu(0) \le c-\epsilon$. $p \notin F^{-1}[c-\epsilon,
c+\epsilon]$. Together with the preceding theorem we get

@@

  We check that the homotopy lies entirely in $H =
  \{\epsilon-\mu(\xi+2\eta) \le \xi-\eta \le \epsilon \}$. It does,
  indeed, since $\epsilon \ge \xi > \xi-t^2\eta > \xi-\eta \ge
  \epsilon-\mu(\xi+2\eta) \ge \epsilon-\mu(\xi+2t\eta)$.

  In region (2), we project this part into $M^{c-\epsilon}$, in the
  $\eta$-direction as the in case (1). Let the following denote the projection:
  \begin{equation}
    (u^1, .., u^n) \mapsto
    (u^1, .., u^\lambda, cu^{\lambda+1}, .., cu^n) \in M^{c-\epsilon}
  \end{equation}

  Then it's easy to calculate $c =
  (\frac{\xi-\epsilon}{\eta})^{\frac{1}{2}}$. Then the homotopy is simply the
  line along the projection:
  \begin{equation}
    r_t: (t, (u^1,..,u^n)) \mapsto
    (u^1, .., u^\lambda, s_tu^{\lambda+1}, .., s_tu^n),
  \end{equation}
  where
  \begin{equation}
    s_t = t + (1-t)c.
  \end{equation}

  Again we check that the homotopy lies entirely in $H$. This is true, since
  $s_t \in [0,1]$. More precisely, $\xi-s_t^2\eta \ge \xi-\eta \ge
  \epsilon-\mu$.

@@

\begin{rmk}
  More generally suppose that there are $k$ non-degenerate critical points
  $p_1$, .., $p_k$ with indices $\lambda_1$, .., $\lambda_k$ in
  $f^{-1}(c)$. Then a similar proof shows that $M^{c+\epsilon}$ has the homotopy
  type of $M^{c-\epsilon}\cup e^{\lambda_1}\cup .. \cup e^{\lambda_k}$.
\end{rmk}

\begin{lem}[Whitehead]
  Let $\phi_0$ and $\phi_1$ be homotopic maps from the sphere $\dot{e}^\lambda$
  to $X$. Then the identity map of $X$ extends to a homotopy equivalence
  \begin{equation}
    k: X \cup_{\phi_0} e^\lambda \to X \cup_{\phi_1} e^\lambda
  \end{equation}
\end{lem}

@@

\begin{lem}[Index theorem for $L_p$]
  The index of $L_p$ at a non-degenerate critical point $q \in M$ is equal to
  the number of focal points of (M, q) which lie on the segment from $q$ to $p$;
  each focal point being counted with its multiplicity.
\end{lem}

@@

\begin{asser}
  The focal points of $(M, q)$ along any normal lines $\mathbb{l}$ ocur in pairs
  which are situated symmetrically about $q$.
\end{asser}

\begin{proof}

>>>>

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath}               % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{amssymb}
\usepackage[all]{xy}

@@

\begin{abstract}
  This report covers roughly the part I of J. Milnor's book\cite{Mil}. The main
  result is the theorem: every smooth manifoldhas the same homotopy type of a
  CW-complex. We prove this theorem by developing the concept of non-degenerate
  smooth functions, their index at critical points and the correspondence
  between the index and the dimension of attaching cells. After the main
  theorem, the existence of non-degenerate smooth functions on manifolds is
  established, and as an application, a proof of the Lefschetz hyperplane
  theorem is given.
\end{abstract}

@@

  We may assume $h_{ij}$ is symmetric, since we can write $\bar{h}_{ij} =
  \frac{1}{2}(h_{ij}+h_{ji})$ and get $f = \sum_{i,j}
  x^ix^j\bar{h}_{ij}$. Moreover, $\bar{h}_{ij}(0) = \frac{1}{2}\frac{\partial^2
    f}{\partial x^i \partial x^j}(0)$, so it is non-singular.

@@

\begin{thm}
  \label{3.1}
  Suppose $f^{-1}[a,b]$ is compact and contains no critical points of $f$. Then
  $M^a$ is diffeomorphic to $M^b$. Furthermore, $M^a$ is a deformation retract
  of $M^b$, so that the inclusion map $\iota : M^a \to M^b$ is a homotopy
  equivalence.
\end{thm}

@@

  For fixed $q \in M$ consider the function $t \mapsto f(\phi_t(q))$. Then If
  $\phi_t(q) \in f^{-1}[a,b]$, then
  \begin{equation}
    \frac{d f(\phi_t(q))}{dt} = X_{\phi_t(q)}(f) = 
    \langle \rho(grad \:  f), grad \: f \rangle \le 1.
  \end{equation}
  Hence the inequality $f(\phi_t(q)) \le t + f(q)$ holds for all $t \in \RR$. It

  Consider the diffeomorphism $\phi_{b-a} : M \to M$. If $p \in M^b$, let $q =
  \phi_{a-b}(p)$. Then $\phi_{b-a}(q) = p$ and $f(q) = f(\phi_{a-b}(p)) \le
  a-b+f(p) \le a$. So $\phi_{b-a}$ maps $M^a$ onto $M^b$. It follows that
  $\phi_{b-a}$ maps $M^a$ diffeomorphically onto $M^b$. This prove the first
  part of the theorem.

@@

\begin{thm}
  \label{3.2}
  Let $f: M \to \RR$ be a smooth function and let $p$ be a non-degenerate
  critical point with index $\lambda$. Setting $f(p) = c$, suppose that $p$ is
  the only critical point in $f^{-1}[c-\epsilon, c+\epsilon]$, for some
  $\epsilon > 0$. Then for all $\epsilon$ sufficiently small, the set
  $M^{c-\epsilon}$ has the same homotopy type of $M^{c+\epsilon}$ with a
  $\lambda$-cell attached.
\end{thm}

@@

For example, we can define $\mu$ as follows. Let $\mu = \epsilon'
\frac{\int_x^{2\epsilon} \alpha(t)\alpha(2\epsilon-t)dt}{\int_0^{2\epsilon}
  \alpha(t)\alpha(2\epsilon-t)}dt$, where $2\epsilon > \epsilon' > \epsilon$ and
\begin{equation}
  \alpha(x) =
  \begin{cases}
    0 & x \le 0 \\
    e^{-1/x} & x > 0
  \end{cases}
\end{equation}
is smooth. Then $0 \ge \mu'(r) > -\frac{\epsilon'}{2\epsilon} > -1$.

@@

\begin{proof}
  \begin{equation}
    \nonumber
    \begin{aligned}
      F &= c - \xi + \eta - \mu(\xi+2\eta) \\
      \frac{\partial F}{\partial \xi} &= -1 - \mu'(\xi+2\eta) < 0 \\
      \frac{\partial F}{\partial \eta} &= 1 - 2\mu'(\xi+2\eta) \ge 1 \\
      dF &= \frac{\partial F}{\partial \xi}d\xi+\frac{\partial F}{\partial
        \eta}d\eta
    \end{aligned}
  \end{equation}
  Since the 1-forms $d\xi$ and $d\eta$ are simultaneously 0 only when both
  $\xi$ and $\eta$ equal 0, it follows that $F$ has no critical point in $U$
  other than $p$.
\end{proof}

Now, since $F \le f$, if $q \in F^{-1}[c-\epsilon, c+\epsilon]$, then
$c-\epsilon \le f(q)$. But $q \in F^{-1}(-\infty, c+\epsilon] = f^{-1}(-\infty,
c+\epsilon]$, $f(q) \le c+ \epsilon$. Therefore $q \in f^{-1}(-\infty,
c+\epsilon]$. This proves that $F^{-1}[c-\epsilon, c+\epsilon] \subset
f^{-1}[c-\epsilon, c+\epsilon]$. Being a closed subset of a compact set in a
locally compact Hausdorff space, $F^{-1}[c-\epsilon, c+\epsilon]$ is compact.
Note that $F(p) = c- \mu(0) < c-\epsilon$. $p \notin F^{-1}[c-\epsilon,
c+\epsilon]$. Together with the preceding theorem we get

@@

  We check that the homotopy lies entirely in $H =
  \{\epsilon-\mu(\xi+2\eta) \le \xi-\eta \le \epsilon \}$. It does,
  indeed, since $\epsilon \ge \xi \ge \xi-t^2\eta \ge \xi-\eta \ge
  \epsilon-\mu(\xi+2\eta) \ge \epsilon-\mu(\xi+2t^2\eta)$.

  In region (2), we project this part into $M^{c-\epsilon}$, in the
  $\eta$-direction as the in case (1). Let the following denote the projection:
  \begin{equation}
    (u^1, .., u^n) \mapsto
    (u^1, .., u^\lambda, ku^{\lambda+1}, .., ku^n) \in M^{c-\epsilon}
  \end{equation}

  Then it's easy to calculate $k =
  (\frac{\xi-\epsilon}{\eta})^{\frac{1}{2}}$. Then the homotopy is simply the
  line along the projection:
  \begin{equation}
    r_t: (t, (u^1,..,u^n)) \mapsto
    (u^1, .., u^\lambda, s_tu^{\lambda+1}, .., s_tu^n),
  \end{equation}
  where
  \begin{equation}
    s_t = t + (1-t)k.
  \end{equation}

  Again we check that the homotopy lies entirely in $H$. We first note that $s_t
  \in [0,1]$. Then $\xi-s_t^2\eta \ge \xi-\eta \ge \epsilon-\mu$. Also,

  \begin{equation}
    \begin{aligned}
      \xi-s_t^2\eta &= \xi -
      (t^2+2t(1-t)(\frac{\xi-\epsilon}{\eta})^{1/2}+(1-t)^2\frac{\xi-\epsilon}{\eta})\eta
      \\
      &=
      \xi-t^2\eta-2(t-t^2)\sqrt{\eta(\xi-\epsilon)}-(1-t)^2\epsilon+(1-t)^2\epsilon
      \\
      &\le
      \xi-t^2(\epsilon-\xi)-(2t-2t^2)(\xi-\epsilon)-(1-t)^2\xi+(1-t)^2\epsilon
      \\
      &= \epsilon,
    \end{aligned}
  \end{equation}
  where we use the condition $\xi-\eta \le \epsilon$.

@@

\begin{rmk}
  \label{3.3}
  More generally suppose that there are $k$ non-degenerate critical points
  $p_1$, .., $p_k$ with indices $\lambda_1$, .., $\lambda_k$ in
  $f^{-1}(c)$. Then a similar proof shows that $M^{c+\epsilon}$ has the homotopy
  type of $M^{c-\epsilon}\cup e^{\lambda_1}\cup .. \cup e^{\lambda_k}$.
\end{rmk}

\begin{thm}
  \label{main}
  If f is a differentiable function on a manifold $M$ with no degenerate
  critical points, and if each $M^a$ is compact, then $M$ has the homotopy type
  of a CW-complex with one cell of dimension $\lambda$ for each critical point
  of index $\lambda$.
\end{thm}

The proof will be based on two lemmas concerning a topological space $X$ with a
cell attached. The first lemma says that if two attaching methods are homotopic
to each other, then the two resulting spaces are homotopy equivalent. The second
lemma shows that if the base space is changed by a map preserving its homotopy
type, then then the two attaching spaces are homotopic equivalent in a natural
way.

\begin{lem}[Whitehead]
  \label{whitehead}
  Let $\phi_0$ and $\phi_1$ be homotopic maps from the sphere $\dot{e}^\lambda$
  to $X$. Then the identity map of $X$ extends to a homotopy equivalence
  \begin{equation}
    k: X \cup_{\phi_0} e^\lambda \to X \cup_{\phi_1} e^\lambda
  \end{equation}
\end{lem}

\begin{proof}
  Let $\phi_t$ be the homotopy between $\phi_0$ and $\phi_1$. Let $u$ denote a
  point on $\dot{e}^\lambda$. We define $k$ by the following formulas:
  \begin{equation}
    \begin{array}{lcl}
      k(x) = x, & x \in X, \\
      k(tu) = 2tu & t \in [0,\frac{1}{2}], \\
      k(tu) = \phi_{2-2t}(u) & t \in [\frac{1}{2},1].
    \end{array}
  \end{equation}

  Notice that $k(\frac{1}{2}u) = u = \phi_1(u)$.

  Define $l : X \cup_{\phi_1} e^\lambda \to X \cup_{\phi_0} e^\lambda$ in a
  similar way as $k$:

  \begin{equation}
    \begin{array}{lcl}
      k(x) = x, & x \in X, \\
      k(tu) = 2tu & t \in [0,\frac{1}{2}], \\
      k(tu) = \phi_{2t-1}(u) & t \in [\frac{1}{2},1].
    \end{array}
  \end{equation}

  Then $lk$ is defined as the following formulas

  \begin{equation}
    \begin{array}{lcl}
      lk(x) = x, & x \in X, \\
      k(tu) = 4tu & t \in [0,\frac{1}{4}], \\
      k(tu) = \phi_{4t-1}(u) & t \in [\frac{1}{4},\frac{1}{2}], \\
      k(tu) = \phi_{2-2t}(u) & t \in [\frac{1}{1},1].
    \end{array}
  \end{equation}

  Hence there is a homotopy between $lk$ and identity

  \begin{equation}
    \begin{array}{lcl}
      H(s, x) = x, & s \in [0,1], x \in X,\\
      H(s, tu) = (4-3s)tu & t \in [0,\frac{1}{4-3s}], \\
      H(s, tu) = \phi_{4t-1}(u) & t \in [\frac{1}{4-3s},\frac{2-s}{4-3s}], \\
      H(s, tu) = \phi_{\frac{4-3s}{4}(2-2t)}(u) & t \in [\frac{2-s}{4-3s},1].
    \end{array}
  \end{equation}

  $H$ is continuous on a finite number of closed sets of $[0,1]\times[0,1]$ and
  well-defined, so $H$ is continuous.

  In a similar fashion, we can construct a homotopy from $kl$ to identity

  \begin{equation}
    \begin{array}{lcl}
      \bar{H}(s, x) = x, & s \in [0,1], x \in X,\\
      \bar{H}(s, tu) = (4-3s)tu & t \in [0,\frac{1}{4-3s}], \\
      \bar{H}(s, tu) = \phi_{2-(4-3s)t}(u) &
      t \in [\frac{1}{4-3s},\frac{2-s}{4-3s}], \\
      \bar{H}(s, tu) = \phi_{\frac{(4-3s)t}{2}+\frac{3}{2}s-1}(u) &
      t \in [\frac{2-s}{4-3s},1].
    \end{array}
  \end{equation}

\begin{lem}
  \label{3.7}
  Let $\phi : \dot{e}^\lambda \to X$ be an attaching map. Any homotopy
  equivalence $f: X \to Y$ extends to a homotopy equivalence
  \begin{equation}
    F: X \cup_\phi e^\lambda \to Y \cup_{f\phi} e^\lambda
  \end{equation}
\end{lem}

\begin{proof}
  Define $F|_X = f$, $F|_{e^\lambda} = identity$. Let $g$ be the homotopy
  inverse of $f$. Define $G: Y \cup_{f\phi} e^\lambda \to X \cup_{gf\phi}
  e^\lambda$ by $G|_Y = g$, $G|_{e^\lambda} = identity$. Since $gf\phi$ is
  homotopic to $\phi$, by \ref{whitehead} there is a homotopy equivalence
  $G: X \cup_{gf\phi} e^\lambda \to X \cup_{\phi} e^\lambda$. We want to show
  that $kG$ is a homotopic inverse of $F$. We first write down the composition
  $kGF$ in the following formulas

  \begin{equation}
    \begin{array}{lcl}
      kGF(x) = x, & x \in X, \\
      kGF(tu) = 2tu & t \in [0,\frac{1}{2}], \\
      kGF(tu) = h_{2-2t}(u) & t \in [\frac{1}{2},1]
    \end{array}
  \end{equation}
  where $h_t$ is a homotopy between $gf$ and identity. $kGF$ is homotopic to
  identity by the following formulas

  \begin{equation}
    \begin{array}{lcl}
      H(s, x) = h_s(x), & s \in [0,1], x \in X,\\
      H(s, tu) = (2-s)tu & t \in [0,\frac{1}{2}], \\
      H(s, tu) = h_{2-(2-s)t}(tu) & t \in [\frac{1}{2},1]
    \end{array}
  \end{equation}

  Thus $F$ has a left inverse. Note that $g$ and the induced $G$ also satisfy
  the condition of the lemma. So $G$ also has a left inverse.

  It is elementary that if $F$ has both a left inverse $L$ and a right inverse
  $R$, then $F$ is a homotopy equivalence and both $L$ and $R$ are two sided
  inverses.

  Now we have $k(GF) \sim id$. Since $k$ is a homotopy equivalence, $GFk \sim
  id$. Thus $G$ has a right inverse. Then $G$ is also a homotopy equivalence and
  $FkG \sim id$. Thus $F$ also has a right inverse. This proves that $F$ is a
  homotopy equivalence.
\end{proof}

We now go back to the proof of theorem \ref{main}:

\begin{proof}
  Let $c_1 < c_2 < c_3 < ..$ be the critical values of $f: M \to \RR$. The
  sequence ${c_i}$ has no cluster points, for each $M^a$ is compact \footnote{If
    $c$ is a cluster point of ${c_i}$,then there exists a sequence ${p_j}$ in M
    such that $f(p_j) = c_{i_j}$ and $\lim_{j \to \infty} p_j = p$ If p is not a
    critical point, then there is a neighborhood $U$ of $p$ such that each point
    of $U$ is not a critical point. But this is a contradiction, since $p$ is
    the limit of the sequence ${p_j}$.}. We first prove by induction that each
  level set $M^a$ has the homotopy type of a CW-complex. If $a < c_1$, then $M^a
  = \emptyset$. Suppose $c_{i-1} < a < c_{i} =: c$. By theorems \ref{3.1},
  \ref{3.2} and \ref{3.3}, $M^{c+\epsilon} = M^a \cup_{\phi_1} e^{\lambda_1}
  \cup .. \cup_{\phi_{j(c)}} e^{\lambda_{j(c)}}$, for sufficiently small
  $\epsilon$. By induction hypothesis, there is a homotopy equivalence $h : \to
  K$.By lemma \ref{3.7}, $h$ induces a homotopy equivalence between
  $M^{c+\epsilon} = M^a \cup_{\phi_1} e^{\lambda_1} \cup .. \cup_{\phi_{j(c)}}
  e^{\lambda_{j(c)}}$ and $K \cup_{h\circ\phi_1} e^{\lambda_1} \cup
  .. \cup_{\circ\phi_{j(c)}} e^{\lambda_{j(c)}}$

  Each $h \circ \phi_j$ is homotopic by cellular approximation to
  \begin{equation}
    \psi_j: \dot{e}^{\lambda_j} \to K^{\lambda_j-1}.
  \end{equation}

  By lemma \ref{whitehead}, replacing the attaching map does not
  change the homotopy type. Thus we have shown that $M^{c+\epsilon}$
  has the homotopy type of a CW-complex.

  If M is compact, then $M = M^a$ for some $a \in \RR$ and the proof is done. If
  M is not compact, but all critical points lie in $M^a$, a proof similar to
  \ref{3.1} shows that $M^a$ is a deformation retract of $M$.

  If there are infinitely many critical points, then we can construct an
  infinite sequence of homotopy equivalences
  \begin{equation*}
    \xymatrix@C+0em@R+0em{
      M_{a_1} \ar[d]_{g_1} \ar@{^(->}[r] &
      M_{a_2} \ar[d]_{g_2} \ar@{^(->}[r]&\cdots \\
      K_1 \ar@{^(->}[r] & K_2 \ar@{^(->}[r] &\cdots
    }
  \end{equation*}
  each extending the previous one. Let $K = \cup_{i=1}^{\infty}K_i$ have the
  direct limit topology. We show that $K$ is a CW-complex. Let $C$ be a closed
  set in $K$. Then by definition, $C\cap K_i$ is closed in $K_i$ for all
  $i$. This means that $C\cap e_{ij}$ is closed in $e_{ij}$ for all $e_{ij} \in
  K_i$ for all $i$. Thus $C \cap e$ is closed for all cell in $K$. This proves
  that $K$ has the weak topology.

  Now let $C$ be a compact set in $K$. Let $I=\{ p_i | p_i \in C \cap
  e_i\}$, where $e_i$ ranges over all cells in $K$. Then $I$ is
  closed, since the intersection of $I$ and each cell is a point. $I$
  is descrete, since each $p \in I$ lies in a cell and each cell is
  closure-finite. This proves that $K$ has the closure-finite
  property.

  Let $g: M \to K$ be the limit map. We need Whitehead's theorem to conclude
  that $g$ is a homotopy equivalence:

  \begin{thm}
    If $M$ and $K$ are both dominated by CW-complexes, then any map $M \to K$
    which induces isomorphisms of homotopy groups is a homotopy equivalence.
  \end{thm}

  \begin{defn}\cite{Hat}
    A space $Y$ is dominated by a space $X$ if there are maps
    \begin{equation}
      \xymatrix{
        Y \ar[r]^i & X \ar[r]^r & Y
      }
    \end{equation}
    such that $ri \sim 1_Y$.
  \end{defn}

  $M$ can be dominated by a CW-complex in the following way. Embed $M$ into
  $\RR^n$. Let $N$ be a tubular neighborhood of $M$. Triangulate $\RR^n$ so that
  the simplex intersecting $M$ lies entirely in $N$. Collecting all the
  simplexes and their faces, we get a CW-complex $K$. $M \subset W$. The
  retraction of the tubular neighborhood into $M$ restricting on $W$ is a
  retraction on $W$ into $M$. This shows that $M$ is dominated by a CW-complex.

  It remains to show that $g_*$ is an isomorphism of each $\pi_n(M)$ and
  $\pi_n(K)$. We prove the onto-ness first. Let $f_i$ denote the homotopy
  inverse of $g_i$. Let $[\phi] \in \pi_n(K)$. Since $\phi(S^n)$ is compact,
  $\phi(S^n)$ lies in some $K_i$. Since $g_i\circ f_i \circ \phi$ is homotopic
  to $\phi$ in $K_i \subset K$, $g_*[f_i\circ\phi] = [\phi]$. Next we prove that
  $g_*$ is one-to-one. Suppose there exist $\phi_0, \phi_1 : S^n \to K$ such
  that $g\circ\phi_0 \sim g\circ\phi_1$ in $K$. Denote the homotopy as
  $H$. Since $H$ is continuous and $[0,1]\times S^n$ is compact, the homotopy
  lies in a compact set of $K$, thus in $K_i$ for some $i$. Therefore $f_i\circ
  g_i\circ\phi_0 \sim f_i\circ g_i\phi_1$ in $M^{a_i}$. Hence $f_i\circ
  \phi_0 \sim \phi_1$ in $M$.

  The theorem follows.

@@

\begin{lem}[Index theorem for $L_p$]
  \label{index}
  \label{index_of_Lp}
  The index of $L_p$ at a non-degenerate critical point $q \in M$ is equal to
  the number of focal points of (M, q) which lie on the segment from $q$ to $p$;
  each focal point being counted with its multiplicity.
\end{lem}

@@

\begin{asser}
  The focal points of $(M, q)$ along any normal lines occur in pairs
  which are situated symmetrically about $q$.
\end{asser}

\begin{proof}
  Let $(z^1, .., z^k)$ be coordinates centered at $q$ such that $g_{ij}(q) =
  \delta_{ij}$\footnote{
    Let $E_1, .., E_k$ be an orthornormal basis in $T_qM$. Let $(u_1, .., u_k)$
    be a local coordinates at $q$. Then $E_i = a_{ji}\frac{\partial w}{\partial
      u^j}(q)$, where the matrix $(a_{ij})$ is non singular. Define $(v_1, ..,
    v_k)$ by $u_i = a_{ij}v^j$. Then $\frac{\partial w}{\partial v_k} =
    \frac{\partial w}{\partial u^j}\frac{\partial u^j}{\partial v^i} = E_i$.}
  . The embedding determines analytic coordinate functions
  $w_\alpha(z^1, .., z^k)$, $\alpha = 1, .., n$.  In $\CC^n$, we consider the
  Hermitian inner product $ v \cdot w = \sum_i v_i\bar{w}_i$.

  Let $v$ be a unit vector in $\CC^n$, orthogonal to $M$ at $q$. We consider the
  quadratic forms

  \begin{equation}
    Q(z^1, .., z^k) = 
    \sum_{i,j} v \cdot \frac{\partial^2 w}{\partial z^i \partial z^j}z^iz^j
  \end{equation}
  and

  \begin{equation}
    \label{QQ}
    Q'(x^1, .., x^k, y^1, .., y^k) = Re\{Q(z^1, .., z^k)\}
  \end{equation}
  where $z^j = x^j+iy^j$.

  We have

  \begin{equation}
    \begin{aligned}
      \frac{\partial^2 w}{\partial z^i \partial z^j} &=
      \frac{1}{4}\left(
        \frac{\partial}{\partial x^i}-i\frac{\partial}{\partial
          y^i}\right)\left(\frac{\partial}{\partial
          x^j}-i\frac{\partial}{\partial y^j}
      \right)w \\
      &= \frac{\partial^2 w}{\partial x^i \partial x^j} =
      -\frac{\partial^2 w}{\partial y^i \partial y^j} = 
      -i\frac{\partial^2 w}{\partial x^i \partial y^j} = 
      -i\frac{\partial^2 w}{\partial y^i \partial x^j}
    \end{aligned}
  \end{equation}
  Substituting the above relations into \eqref{QQ}, we find that $Q'$ is exactly
  the second fundamental form of $M$ at $q$ in $\RR^{2n}$.  Recall that the
  focal points are exactly $q+K_i^{-1}v$, where $K_i$ are the principal
  curvatures. From the previous assertion, we see that $-K_i$ is also a
  principal curvature. Thus the focal points lie symmetrically about $q$ on the
  line passing through $q$ in the $v$ direction.
\end{proof}

Now we go back to the proof of the theorem. Choose $p$ in $\CC^n$ such that
$L_p$ is non-degenerate. Since $M$ is a closed subset of $\CC^n$, each $M^a =
L_p^{-1}[0, a]$ is compact. For each $q$ there are at most $2k$ focal points
that lie on line $\stackrel{\leftrightarrow}{pq}$. Since they are distributed
about $q$, there are at most $k$ focal points that lie on the line segment from
$q$ to $p$. By lemma \ref{index}, the index of $L_p$ at $q$ is at most $k$. By
theorem \ref{main}, $M$ has the homotopy type of a CW-complex of dimension at
most $k$. This completes the proof.

\begin{cor}
  If $M \subset \CC^n$ is a non-singular affine algebraic variety in $\CC^n$
  with real dimension $2k$, then
  \begin{equation}
    H_i(M; \ZZ) = 0
  \end{equation}
  for $i>k$.
\end{cor}

\begin{proof}
  $M$ has the homotopy type of a CW-complex of dimension $\le k$. Since $M$ has
  no cells of dimension $ > k$, the $i$th cellular chain complex is $0$.
\end{proof}

\begin{cor}[Lefschetz]
  Let $V$ be an algebraic variety of complex dimension $k$ which lies in the
  complex projective space $\CC P_n$. Let $P$ be a hyperplane in $\CC P_n$ which
  contains the singular points (if any) of $V$. Then the inclusion map
  \begin{equation}
    V\cup P \to V
  \end{equation}
  induces isomorphisms of homology groups in dimensions less than
  $k-1$. Furthermore, the induced homomorphism
  \begin{equation}
    H_k(V\cup P;\ZZ) \to H_{k-1}(V;\ZZ)
  \end{equation}
  is onto.
\end{cor}

\begin{proof}
  The short exact sequence
  \begin{equation}
    \xymatrix{
      0 \ar[r] &
      S_{\cdot}(V\cap P;\ZZ) \ar[r] &
      S_{\cdot}(V;\ZZ) \ar[r] &
      S_{\cdot}(V, V\cap P;\ZZ) \ar[r] & 
      0
    }
  \end{equation}
  induces the long exact sequence
  \begin{equation}
    \xymatrix{
      \cdots \ar[r] &
      H_{r+1}(V, V\cap P;\ZZ) \ar[r] &
      H_{r}(V\cap P;\ZZ) \ar[r] &
      H_{r}(V,;\ZZ) \ar[r] & 
      H_{r}(V, V\cap P;\ZZ) \ar[r] &
      \cdots
    }
  \end{equation}

  Therefore it suffices to show that $H_r(V, V\cap P; \ZZ) = 0$ for $r \le
  k-1$. The Lefschetz duality theorem asserts that
  \begin{equation}
    H_r(V,V\cap P;\ZZ) \cong H^{2k-r}(V-(V\cap P);\ZZ).
  \end{equation}

  Since $V-(V\cap P)$ is a complex analytic manifold embedded as a closed set in
  $\CC^n$, it has the homotopy type of a CW-complex of dimension $\le
  k$. Therefore, by the previous corollary, $H_{2k-r}(V-(V\cap P);\ZZ) = 0$ for
  $r \le k-1$.
\end{proof}

\begin{thebibliography}{9}                                                                                                
%\bibitem{Car} \textsc{Carmo, M. do},\ \textit{Riemannian Geometry}, Birkh\"auser, 1992.

\bibitem{Mil} \textsc{Milnor, J.},\ \textit{Morse Theory}, Princeton University
  Press, 1969.

\bibitem{Hat} \textsc{Hatcher, A.},\ \textit{Algebraic Topology}, Cambridge
  University Press, 2002.

\bibitem{Spa} \textsc{Spanier, E.},\ \textit{Algebraic Topology}, McGraw-Hill,
  1966.

\end{thebibliography}
